---
title: "Polling Simulation: Sampling and the Central Limit Theorem"
---

In this activity, you'll use data from the 2024 Cooperative Election Study (CES) — a large-scale national survey conducted during the 2024 election — to simulate what happens when a news organization tries to estimate public opinion from a poll sample. The full dataset has nearly 55,000 respondents, which we'll treat as our population.

Your goal is to see firsthand how sample size affects the accuracy of estimates, and what the distribution of sample estimates actually looks like — the core idea behind the Central Limit Theorem.

### Task 1: Load the tidyverse

```{r}
library(tidyverse)
```

### Task 2: Load the CES data

Load the CES 2024 religion dataset from this URL and save it to a dataframe called `ces`:

`https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/ces_2024_religion.csv`

```{r}
ces <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/ces_2024_religion.csv")
```

### Task 3: Calculate the population proportion

We'll focus on one question: what share of American adults identify as religiously unaffiliated (atheist, agnostic, or nothing in particular)? This has been a major story in American religion journalism for years.

First, we'll create a new column called `unaffiliated` that is `TRUE` if a respondent's `trad` value is `"Atheist"`, `"Agnostic"`, or `"Nothing in Particular"`, and `FALSE` otherwise. Then calculate the proportion of the full dataset that is unaffiliated. Save this value as `population_proportion`.

```{r}
ces <- ces |>
  mutate(unaffiliated = trad %in% c("Atheist", "Agnostic", "Nothing in Particular"))

proportion <- mean(ces$unaffiliated)
```

That population proportion is the "true" population value that any poll is trying to estimate. Write it down — you'll be comparing everything else to it.

### Task 4: Take a single small sample

Use `set.seed(42)` and then sample 50 rows from `ces`. Save it as `sample_50`. Calculate the proportion of unaffiliated respondents in your sample.

```{r}
set.seed(42)
sample_50 <- ces |> sample_n(50)
mean(sample_50$unaffiliated)
```

How close is your sample estimate to the original proportion? Is it surprising?

The estimate is pretty close to the original proportion but it still a little far off but not by much. It is pretty surprising how close it is to the original proportion given how the sample size was on the smaller size. 

### Task 5: Take a single larger sample

Now sample 500 rows. Set a different seed value, grab the sample and save it as `sample_500`. Calculate the proportion of unaffiliated respondents in this sample.

```{r}
set.seed(405)
sample_50 <- ces |> sample_n(500)
mean(sample_50$unaffiliated)
```

How did the estimate change compared to the sample of 50? What does this suggest about sample size?

The estimate became closer to the original proportion. This suggests that the higher the sample size the more accurate your results are. 

### Task 6: Simulate many polls

A single sample can get lucky or unlucky. To understand how polling actually works, we need to see the *distribution* of estimates across many samples. The code below draws 1,000 samples of size 50, calculates the proportion of unaffiliated respondents in each, and stores the results.

```{r}
set.seed(405)

simulate_polls <- function(data, sample_size, n_simulations = 1000) {
  proportions <- replicate(n_simulations, {
    s <- data |> sample_n(sample_size)
    mean(s$unaffiliated)
  })
  tibble(proportion = proportions, sample_size = sample_size)
}

simulations_50 <- simulate_polls(ces, 50)
simulations_500 <- simulate_polls(ces, 500)
```

### Task 7: Visualize the simulation results

Create two histograms side by side — one for the 1,000 samples of size 50 and one for the 1,000 samples of size 500. Add a vertical line at `proportion` to each so you can see how well the samples are centering on the true value.

```{r}
simulations_all <- bind_rows(simulations_50, simulations_500) |> 
  mutate(sample_size = paste("n =", sample_size))

ggplot(simulations_all, aes(x = proportion)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "white") +
  geom_vline(xintercept = proportion, color = "red", linetype = "dashed", linewidth = 1) +
  facet_wrap(~sample_size) +
  labs(
    title = "Distribution of Sample Estimates: Share Religiously Unaffiliated",
    subtitle = "Red line = true population proportion. Each histogram shows 1,000 simulated polls.",
    x = "Sample Proportion",
    y = "Count",
    caption = "Source: 2024 Cooperative Election Study"
  ) +
  theme_minimal()
```

### Task 8: Reflect on what you see

**What shape do both histograms have?** Describe them in plain language. Does it surprise you that this shape emerges from a variable that is simply TRUE or FALSE for each respondent?

The first histogram has a pretty normal distribution but it is pretty spaced out compared to the second histogram. The second histogram looks to have more data and taller peaks in the middle/less spaced out. I wouldn't confidently say any of them are clearly skewed in any direction. This shape doesn't surprise me since the variable is just true or false it makes sense that neither histogram is skewed one way or another and both of them are farily normal. 

**Compare the two histograms.** What is the most visible difference between the n=50 and n=500 distributions? What does that difference tell you about what happens when a poll uses a larger sample?

The most visible difference between the two histograms is that the n=50 histogram is more spread out and has lower peaks while the n=500 histogram has more data and taller peaks and also looks to be more crammed together. This tells me that when you poll a larger sample it may have a similar distribution but with larger amounts of data. 

**Both distributions are centered near the red line.** What does it mean that 1,000 different random samples — each one drawn independently — tend to cluster around the same value? What would it mean for journalism if that *weren't* the case?

1,000 different random samples clustering around the same value means that the samples are accurate with the more data showing an accurate sample. If this weren't the case that would mean that there are more outliers within the data. This would make a good story showing the different outliers compared to the rest of data. As a journalist I would also look into this and see if there is any pattern that shows why there are multiple outliers in such a large sample size. 

**A reporter tells you their story is based on a poll of 50 people.** Based on what you've seen in the simulation, what questions would you want to ask before trusting that number? You don't need to use any technical terms — just describe what you observed.

Throughout the activities today I observed that the larger your sample size is the more accurate data you will collect. If a reporter tells me their story is only based off of 50 people I would ask what the story is about. Generally 50 people is on the smaller side and I would suggest maybe sending a digital survey rather than in person if that makes it easier for the reporter instead of interviewing everyone individually. I would also suggest the reporter to look for at least 100 if not much much higher in order to ensure that reporter collects accurate data. This is also proved by the central limit theorem. 

When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
