---
title: "Sampling"
---

Follow along with the demo to learn how to sample data from a population using R and the Tidyverse. You should get started by loading the tidyverse in a new R block.

### Task 1: Load the tidyverse

```{r}
library(tidyverse)
```


### Task 2: Load 1,000 rows of data with the ages of people from 18-80

```{r}
set.seed(1234)

population <- tibble (
  age = sample(18:80, 1000, replace = TRUE)
)
```


### Task 3: Sample 100 rows of data from the dataset

```{r}
sample100 <- population |>
  sample_n(100)
```


### Task 4: Calculate the mean age and standard deviation of the sample

```{r}
sample100|>
  summarize(
  mean_age=mean(age),
    sd_age = sd(age)
  )
```



### Task 5: Calculate the mean age and standard deviation of the population
```{r}
population |>
  summarize(
  mean_age=mean(age),
    sd_age = sd(age)
  )

```


### Task 6: Compare the sample and population statistics

Compare the means and standard deviations. Are they similar? Go back and re-run the sampling and statistics calculations with a smaller sample size. How do the results change?

Yes the mean and standard deviations for the sample and population statistics are only a few decimals off from each other. When I re-ran task #3 I was given a new mean and standard deviation but it was around the same size as the first time I ran it. Every time I ran both the sampling and the population statistics the mean was around 48-52 and the standard deviation was around 16-19. As seen in class if I were to make the sample size a lot smaller the standard deviation would decrease more as it gets smaller but will always be in the same ball park as the rest of them. 

When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
