---
title: "HW2 Prince George's County Crime"
---

In this assignment, we will be working with a dataset reporting the number of crimes by month in Prince George's County, Maryland, for every month in 2024. 

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv. You will create a dataframe called `pg_crime` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
pg_crime <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv")
```
### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, are any of the categories more important than others? Why or why not? Do any of them make the job of summarizing the data more difficult? Why or why not?

# Most of the categories fall into categories themselves. From a reporter standpoint I would say that majority of the crime rates seem consistent throughout the year with not much drastic change. If I were writing a story I would focus on a specific time of year in which a category was higher compared to the rest of the year. For example, B & E commerical seems to have a higher rate in December during the holidays compared to the rest of the year, (it was 75 in December but 36 in January, 31 in Feburary, etc.) 

### Task 4: Calculating Averages

Add up the total number of incidents per month for each category using `mutate()`, calling the new column `total_incidents`. Then calculate the average number of incidents per month for each category of crime using the total number of months, calling the new column `average_incidents`. Assign the result to a new dataframe called `pg_crime_averages`.

```{r}
pg_crime_averages <- pg_crime |> mutate(total_incidents = Jan + Feb + Mar + Apr + May + Jun + Jul + Aug + Sep + Oct + Nov + Dec)
```

```{r}
pg_crime_averages2 <- pg_crime_averages |> mutate(average_incidents = total_incidents/12)
```

### Task 5: Describe the Averages

Examine the new dataframe, `pg_crime_averages`, by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the average number of incidents per month for each category. Which categories have the highest average number of incidents per month? Which have the lowest? What does this tell you about the pattern of crime in Prince George's County? Are there categories that you have questions about? Why or why not?

The categories with the most average number of incidents are "accident" and "auto stolen." I feel like this is pretty expected, as car accidents/thefts are nothing new and are pretty common everywhere so I feel like it isn't necessarily newsworthy unless a specific accident or auto theft is drastically worse than others. However what I do find pretty interesting is that the for the total number of incidents for "robbery residential" it comes up as N/A. I'm not sure if that is an issue with the coding or if there are just that many few numbers of incidents of resident robberies in PG County. As a reporter I think that this is something that could be compared to the auto thefts data, and I would ask "Why are auto thefts more common than residential robberies?"  


### Task 6: Testing a Theory

One theory about crime is that it is more likely to occur in the summer months. Test this theory by creating three new columns using `mutate()` that calculate the percentage change compared to the average for each category of crime for the months of June, July, and August. Call the new columns `june_change`, `july_change`, and `august_change`. Assign the result to a new dataframe called `summer_crime_averages`.

```{r}
summer_crime_averages <- pg_crime_averages2 |> mutate(june_change = (Jun - average_incidents) / average_incidents *100, july_change = (Jul - average_incidents) / average_incidents *100, august_change = (Aug - average_incidents) / average_incidents *100)
```

### Task 7: Describe the Summer Crime Averages

Look at the `summer_crime_averages` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change in the average number of incidents per month for each category of crime for the months of June, July, and August. Do you see any patterns? What does this tell you about the theory that crime is more likely to occur in the summer months?

The average number of crimes committed stayed pretty consistent throughout the summer months, while some categories saw a bigger change during one month, the other two summer months showed the opposite or smaller changes. I would argue that the pattern here is that crime rates flucuate throughout the year and there is no true way to tell when crimes will be worse than other.


### Task 8: What's the Story?

Write up a lede and nutgraf of a story focused on the crime data in Prince George's County. It could be from any of the dataframes you've created in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

Auto thefts more commonly result in no outcome, rather than a satisfying conclusion for the car owner. 

Of the roughly 5,000 auto thefts last year, only 10% were returned to their rightful owners. Approximately 400 cars were stolen each month, and an average of 40 were returned. Despite their best efforts, police fail to return the stolen property more often than not, causing victims to lose their belongings.



### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.
