---
title: "HW3 WH Salaries"
---

In this assignment, we will be working with a dataset containing information about the salaries of White House staffers as of July 2026.

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/white_house_salaries_2026.csv. You will create a dataframe called `wh_salaries` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
wh_salaries <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/white_house_salaries_2026.csv")
```
### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, what stands out to you about this data? Are there any salary values that will cause problems for analysis?

Some salaries are way higher or lower compared to others. For example Luke Brown's salary is $65,500 while other employees such as Harry Jung have salaries as high as $197,200. I find this wide range very interesting and as a reporter I would look into the various jobs and what they do to see how it compares to the difference in salaries. I also noticed that other employees such as Tracy Johnson have a 0 salary which completely skews the dataset resulting in inaccuracies throughout the set and calculations.  

### Task 4: Calculate the Mean and Median of the Salaries

Write code to calculate the mean and median of the salaries in the dataset using the `summarize()` function. You don't need to make a new dataframe, just display the results.

```{r}
wh_salaries |>
  summarize(
    mean_salary = mean(SALARY),
    median_salary = median(SALARY)
  )
```

### Task 5: Add in the minimum and maximum salaries

Take the code you just wrote and add in the minimum and maximum salaries (the functions for that are `min()` and `max()`). Your results should show four columns: mean_salary, median_salary, min_salary, max_salary.

```{r}
wh_salaries |>
  summarize(
    mean_salary = mean(SALARY),
    median_salary = median(SALARY),
    min_salary = min(SALARY),
    max_salary = max(SALARY)
  )
```

### Task 6: Describe the Results

Write a couple of sentences describing the results of your analysis. What do the mean and median salaries tell you about the data? What do the minimum and maximum salaries tell you? Are there any outliers in the data?

The mean and median show that most white house employees earn a salary over $100,000 annually. However that data is skewed by some outliers. A few employees are shown to make $0 annually while others earn up to $197,200 annually. While these outliers are helpful, it shows that the mean and median cannot be accurate without taking out the major outliers in the data. 

### Task 7: Story Idea

Based on the data, what is your best story idea? What is the next step you would take to develop that story idea?

Based on the data, my story would be that some employees are shown to have a $0 salary but are not listed as interns or introductory employees, they are just listed as "employees," at the same scale as the other employees. My story would be to highlight these employees and what they do for the White House compared to other employees who make a significant amount more. 
