---
title: "HW8_MandMs"
---

In this assignment, you'll be picking up where you left off with the M&Ms activity. You'll load the data and perform some basic statistical analysis. Follow the steps below to complete the assignment.

### Task 1: Load the tidyverse

```{r}
library(tidyverse)
```

### Task 2: Load the M&Ms dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv
```{r}
candy <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv")
```


### Task 3: Sample 5 rows of data from the dataset
```{r}
sample_5 <- candy |> sample_n(5)
```


### Task 4: Calculate the mean age and standard deviation of the sample, using either the red or black number of M&Ms
```{r}
sample_5 |> summarize(mean = mean (black), sd = sd(black))
```


### Task 5: Calculate the mean age and standard deviation of the population, using the same measure as in Task 4

```{r}
candy |> summarize(mean = mean (black), sd = sd(black))
```

### Task 6: Compare the sample and population statistics

Compare the two sets of means and standard deviations. Are they similar?

Yes both of my means were around 73-74 and both of my standard deviations were around 15-20. There is a slight difference between them but they are generally in the same range. 

### Task 7: Describe the results

Write a few sentences about what you found in the data. What does this tell you about the M&Ms dataset and about the act of random sampling?

Based on the dataset, I found that in order to have more accurate results you need to have a larger sample size for your data. If the sample size is too small it won't accurately represent the population being covered. In this scenario the 5 sample size mean came pretty close to the population but the standard deviation is still slightly off. The larger the sample size the more accurate it will be. 


When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
